---
title: "Input: IEU Open GWAS Project data"
output: rmarkdown::html_vignette
# date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{Input: IEU Open GWAS Project data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

To create a **gwasglue2** `SummarySet-class()` object we first need GWAS summary data, like the ones that can be obtained from [IEU Open GWAS Project](https://gwas.mrcieu.ac.uk/) through the **ieugwasr** package. 

<br>
<br>

## Setup
In addition to **gwasglue2**, we’ll load the following package:

- [**ieugwasr**](https://mrcieu.github.io/ieugwasr/) - R package for accessing and query the IEU GWAS database

```{r setup}
library(gwasglue2)
library(ieugwasr)
```

<br>
<br>

## Functions
* `create_summaryset()` - Creates a `SummarySet-class()` object from a tibble with OpenGWAS summary data.

<br>
<br>

## Example
We will use the `ieugwasr::associations()` function to query the IEU Open GWAS and obtain GWAS summary data for the region `chr5:74132993-75132993` and the trait id `ukb-d-I9_IHD` (coronary heart disease).

 ```{r}
 data <- ieugwasr::associations(variants = "5:74132993-75132993", id = "ukb-d-I9_IHD")
 data
 ```
 
Then we use the **gwasglue2** constructor `create_summaryset()` to create the `SummarySet-class()`. By default, this function reads a tibble with the same columns names as the  output of the **ieugwasr** package.

```{r}
 sumset_opengwas <- create_summaryset(data)
 sumset_opengwas
```

When we call `sumset_opengwas`, it shows a summary of the `SummarySet` object. How to acess the GWAS summary data is explained in the next section.