---
title: "Input: text data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Input: text data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```


To create a **gwasglue2** `SummarySet-class()` object we first need GWAS summary data, like the ones that can be obtained from [IEU Open GWAS Project](https://gwas.mrcieu.ac.uk/) through the **ieugwasr** package. 

<br>
<br>

## Setup
In addition to **gwasglue2**, we’ll load the following package:

- [**ieugwasr**](https://mrcieu.github.io/ieugwasr/) - R package for accessing and query the IEU GWAS database

```{r setup}
library(gwasglue2)
```

<br>
<br>

## Functions
* `create_summaryset()` - Creates a `SummarySet-class()` object from a `data.frame` or `tibble` with GWAS summary data.

<br>
<br>

## Example
We will need a text file with GWAS summary data we are going to study. For example, we have a text file for  IEU Open GWAS the trait id `ukb-d-I9_IHD` (coronary heart disease) and we want to query for the region `chr5:74132993-75132993`.

 ```{r}
 data <- data.table::fread("ukb-d-I9_IHD.txt", header = TRUE)
 data
 ```
 
Then we use the **gwasglue2** constructor `create_summaryset()` to create the `SummarySet-class()`. By default, this function reads a tibble with the same columns names as the  output of the **ieugwasr** package. To change column names, use the `beta_col`, `se_col`, `samplesize_col`, `pvalue_col`, `chr_col`, `position_col`, `rsid_col`, `effect_allele_col`, `other_allele_col`, `eaf_col`, `id_col` arguments. See the documentation for [`create_summaryset`] for more details.


```{r}
 sumset_text <- create_summaryset(data, variants = "5:74132993-75132993")
 sumset_text
```

When we call `sumset_text`, it shows a summary of the `SummarySet` object. How to acess the GWAS summary data is explained in the next section.
                       
