---
title: "Tutorial 4: Meta analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tutorial 4: Meta analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```


```{r setup}
#library(gwasglue2)
devtools::load_all("../") # this was added just for development

```

Meta-analysis is a statistical combination of the results from two or more separate studies. In gwasglue2, we use the fixed-effect model which assumes that one true effect size underlies all the studies in the meta-analysis.

We are going to perform meta-analysis for two different studies of cardiac heart disease (chd), in the HMGCR (3-hydroxy-3-methylglutaryl-CoA reductase) gene region.

Firt, we choose the IEU ids for the chd trait.

```{r}
ids <- c( "ieu-a-7",  "ukb-d-I9_IHD")
```
Then, we obtain the metadata using `ieugwasr::gwasinfo()` for each study and create a metadata object.

```{r}
metadata <- lapply(seq_along(ids), function(i){
  m <- create_metadata(ieugwasr::gwasinfo(ids[i])) 
})
```

In the <dataset> code bellow, we create an harmonised `dataset` object from the summary sets for each study.

```{r dataset, include = TRUE}
#  create dataset 
  hmgcr_chrpos <- "5:74132993-75132993"
  dataset <- lapply(seq_along(ids), function(i){
    # create summarysets
    s <- create_summaryset(ieugwasr::associations(variants = hmgcr_chrpos, id =ids[i]), metadata=metadata[[i]])
  }) %>%
    # create dataset
    create_dataset(., harmonise = TRUE, tolerance = 0.08, action = 1)
```

Finally, we perform the meta-analysis in <meta> to create a new summary set
 
 ```{r meta, include = TRUE}
 meta_chd <- dataset%>%
    meta_analysis(.)
 ```

 