---
title: "hypercoloc"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{hypercoloc}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```
The **hypercoloc** package performs Bayesian multi-trait Colocalization. We modified the package to accept a `DataSet` object as input.

<br>
<br>


## Setup
In addition to **gwasglue2**, we will load the following package:

- [**hyprcoloc**](https://www.nature.com/articles/s41467-020-20885-8) from a github fork

```{r setup}
remotes::install_github('ritarasteiro/hyprcoloc', build_opts = c('--resave-data', '--no-manual'), build_vignettes = TRUE) 
library(gwasglue2)
library(hyprcoloc)  # fork that takes Dataset
```


<br>
<br>


## Example 

**hypercoloc**  takes a `DataSet` object (`dataset`) like the one created in the [Create a dataset](processing-dataset.html) section.

```{r} 
res_dataset <- hyprcoloc(dataset = dataset) 
#hyprcoloc is using gwasglue2 DataSet class object as input
print(res_dataset)
```

More details can be found [here](finemap_coloc1.html)
