---
title: "Merge two or more datasets"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Merge two or more datasets}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

We previously showed how to create a  **gwasglue2** [`DataSet`](processing-dataset.html) object. Here, we will show how to merge two or more `DataSets`.

<br>
<br>

## Setup

```{r setup}
library(gwasglue2)
library(ieugwasr)
```

<br>
<br>

## Functions
*  `merge_datasets()`: Merge a list of two or more `DataSet` objects.


<br>
<br>

## Example
 The function listed above work on a `DataSet` object. WWe will use `DataSet` objects (`dataset1` and `dataset2`) like the ones created in the [Create a dataset](processing-dataset.html) section. By, default, the `merge_datasets()` function will harmonise again the data.

```{r}
datasets <- list(dataset1, dataset2)
merge_dataset <- merge_datasets(datasets)
```

