---
title: "Processing: Setter and getters functions for the SummarySet object"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Processing: Setter and getters functions for the SummarySet object}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

In the [Input section](input.html), we showed how to create a  **gwasglue2** `SummarySet` object. Here, we will show how to set and get information from a `SummarySet` object.
<br>
<br>

## Setup

```{r setup}
library(gwasglue2)
devtools::load_all("../") # this was added just for development purposes
```

<br>
<br>

## Functions
* `getSummaryData()`: Reads the GWAS summary data from the `SummarySet` object.
* `setMetadata()`: See [here](processing-add-metadata.html)
* `getMetadata()`: See [here](processing-add-metadata.html)
* `addToMetadata()`: See [here](processing-add-metadata.html)
* `setAttributes()`: Sets the attributes of the `SummarySet` object. The `type` and `creation` attributes are added by default.
* `getAttributes()`: Gets the attributes of the `SummarySet` object.
* `setShape()`. The `shape` feature is not yet implemented in **gwasglue2**. 
* `getShape()`. The `shape` feature is not yet implemented in **gwasglue2**.
* `dimData()`. Gives the dimensions of the GWAS summary data.


<br>
<br>

## Example
 All the functions listed above work on a `SummarySet` object. We will use one of the `SummarySet` objects (`sumset_opengwas`) created in the [Input section](input.html).

 We can check the the GWAS summary data with `getSummaryData()` and its dimensions using the `dimData()` function. The dimensions of the GWAS summary data are 4,172 rows and 13 columns. 

```{r}
getSummaryData(sumset_opengwas)
dimData(sumset_opengwas)
```

We can also check the metadata associated with the `SummarySet` object using the `getMetadata()` function. More details on how to add metadata to a `SummarySet` object can be found [here](processing-add-metadata.html).

```{r}
getMetadata(sumset_opengwas)
```

We can also check the attributes associated with the `SummarySet` object using the `getAttributes()` function. By default, **gwasglue2** adds the `type` of GWAS summary data in the `SummarySet` object and the `creation` date. It is also possible to add other attributes to the `SummarySet` object using the `setAttributes()` function. 

```{r}
getAttributes(sumset_opengwas)
sumset_opengwas <- setAttributes(sumset_opengwas, author = "rr")
getAttributes(sumset_opengwas)
```


If we want use the `mr_label = "exposure"` or `mr_label = "outcome"`attribute to the `SummarySet` object for future MR analyses, we can do it as follows:


```{r}
sumset_opengwas <- setAttributes(sumset_opengwas, mr_label = "exposure")
getAttributes(sumset_opengwas)
```