---
title: "Processing: Meta analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Processing: Meta analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```


Meta-analysis is a statistical combination of the results from two or more separate studies. In **gwasglue2**, the meta-analysis is performed using the fixed-effect model, which assumes that one true effect size underlies all the studies. The meta-analysis is performed using the `meta_analysis()` function. This function takes a `DataSet` object as input and returns a new `SummarySet` object.

<br>
<br>

## Setup
In addition to **gwasglue2**, we’ll load the following package:

- [**ieugwasr**](https://mrcieu.github.io/ieugwasr/) - R package for accessing IEU GWAS database

```{r setup}
library(gwasglue2)
library(ieugwasr)
devtools::load_all("../") # this was added just for development
```
<br>
<br>

## Functions
* `create_metadata()`: creates a metadata object
* `create_summaryset()`, using the `metadata` argument: creates a `SummarySet` object
* `create_dataset()`: creates a `DataSet` object 
* `meta_analysis()`: performs meta-analysis and returns a `SummarySet` object
 
 <br>
 <br>
 
## Example

We are going to perform meta-analysis for two different studies of cardiac heart disease (chd), in the HMGCR (3-hydroxy-3-methylglutaryl-CoA reductase) gene region (chr5:74132993-75132993). We will use GWAS summary data obtained from IEU OpenGWAS through the **ieugwasr** package (see [here](input-opengwas.html) for details).

Firt, we choose the IEU ids for the chd trait.

```{r}
ids <- c( "ieu-a-7",  "ukb-d-I9_IHD")
```
Then, we obtain the metadata using `ieugwasr::gwasinfo()` for each study and create a metadata object.

```{r}
metadata <- lapply(seq_along(ids), function(i){
  m <- create_metadata(ieugwasr::gwasinfo(ids[i])) 
})
```

In the <dataset> code bellow, we create an harmonised `dataset` object from the summary sets for each study.

```{r dataset, include = TRUE}
#  create dataset 
  hmgcr_chrpos <- "5:74132993-75132993"
  dataset <- lapply(seq_along(ids), function(i){
    # create summarysets
    s <- create_summaryset(ieugwasr::associations(variants = hmgcr_chrpos, id =ids[i]), metadata=metadata[[i]])
  }) %>%
    # create dataset
    create_dataset(., harmonise = TRUE)
```

Finally, we perform the meta-analysis in <meta> to create a new summary set
 
 ```{r meta, include = TRUE}
 meta_chd <- dataset %>%
    meta_analysis()
 ```

 