---
title: "Processing: Setter and getters functions for the DataSet object"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Processing: Setter and getters functions for the DataSet object}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

We previously showed how to create a  **gwasglue2** [`DataSet`](processing-dataset.html) object. Here, we will show how to set and get information from it.
<br>
<br>

## Setup

```{r setup}
library(gwasglue2)
devtools::load_all("../") # this was added just for development purposes
```

<br>
<br>

## Functions
* `getData()`: Reads the GWAS summary data from one `SummarySet` object within the `DataSet`.
* `getSummarySet()`: Returns a `SummarySet` object within the `DataSet`.
* `getLength()`: Number of `SummarySet` objects within the `DataSet`. 
* `setTraitOrg()`: The `trait organisation` feature is not yet implemented in **gwasglue2**.
* `getTraitOrg()`: The `trait organisation` feature is not yet implemented in **gwasglue2**.
* `assertSameShape()`: The `trait organisation` feature is not yet implemented in **gwasglue2**.

<br>
<br>

## Example
 All the functions listed above work on a `DataSet` object. We will use one of the `DataSet` object (`dataset`) created in the [Create a dataset](processing-dataset.html) section.



First, we will look at how many `SummarySet` objects are within the `DataSet`  with the dimData()` function.  We can see there are two.
 ```{r}
getLength(dataset)
```
Next, we can check each of the `SummarySet` objects within the `DataSet` with `getSummarySet()` and retrieve its GWAS summary data with  `getData()`.  These two functions have an `index` argument to choose the `SummarySet` object within the `DataSet` to retrieve.  If the `index` argument is set to `1` it will retrieve the first `SummarySet` object.

```{r}
getSummarySet(dataset,1)
getData(dataset,1)

```
If the `index` argument is set to `2` it will retrieve the second `SummarySet` object.

```{r}
getSummarySet(dataset,2)
getData(dataset,2)

```


