---
title: "Add one or more SummarySet objects to a DataSe`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Add one or more SummarySet objects to a DataSet}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

We previously showed how to create a  **gwasglue2** [`DataSet`](processing-dataset.html) object. Here, we will show how to add two or more `SummarySet` objects to a `DataSet`.

<br>
<br>

## Setup

```{r setup}
library(gwasglue2)
library(ieugwasr)
```

<br>
<br>

## Functions
*  `add_summaryset()` - Add one or more `SummarySet` objects to a `DataSet` object.


<br>
<br>

## Example
The function listed above work on a `DataSet` object and it takes two arguments: `summary_sets` and `dataset` that take a `SummarySet` or a list of two or more `SummarySet` objects, and the `DataSet`, respectively. This function also has harmonisation arguments and by default,  will harmonise again the data.

We will use a `DataSet` object (`dataset`) like the one created in the [Create a dataset](processing-dataset.html) section and  the `SummarySet` objects from the examples in the [Input section](input.html). 


We can add one `SummarySet`:


```{r}
dataset_1sumset_added <- add_summaryset(sumset_opengwas, dataset)
```

Or a list of two or more `SummarySet` objects:
```{r}
summarysets <- list(sumset_opengwas, sumset_vcf)
dataset_2sumset_added <- add_summaryset(summarysets, dataset)
```