---
title: "Processing: liftover"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Processing: liftover}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

**gwasglue2** has a `liftover()` function to remap genomic coordinates from different genome assemblies to the same reference build. This function works on the `SummarySet` object and uses a chain file to map the coordinates. The chain file is a text file that contains the mapping between two genome assemblies. The chain file can be provided by the user or downloaded directly by the  **gwasglue2** `download_chainfile()` or `liftover()` functions.

**gwasglue2** uses the chain files and Human genome nomenclature of [ENSEMBL](https://www.ensembl.org/Homo_sapiens/) ("NCBI34", "NCBI35", "NCBI36", "GRCh37" and "GRCh38").

<br>
<br>
## Setup
In addition to **gwasglue2**, we’ll load the following package:

- [**ieugwasr**](https://mrcieu.github.io/ieugwasr/) - R package for accessing IEU GWAS database

```{r setup}
library(gwasglue2)
library(ieugwasr)
devtools::load_all("../") # this was added just for development
```
<br>
<br>
## Functions
* `create_summaryset()`, using the `build` argument.
* `download_chainfile()`: downloads Human genomes chain files from ENSEMBL.
* `liftover()`: remaps the genomic coordinates from the GWAS summary data on the `SummarySet` to another genome assembly using a chain file.

<br>
<br>

## Example

We will use GWAS summary data obtained from IEU OpenGWAS through the **ieugwasr** package (see [here](input-opengwas.html) for details). The data was built under the "GRCh37" genome assembly and we want to remap it to "GRCh378". 

First, we create a SummarySet object that we name `sumset` providing in the `build` argument the genome assembly used to build the GWAS summary data. We can check it in the metadata.


 ```{r}
 data <- ieugwasr::associations(variants = "5:74132993-75132993", id = "ukb-d-I9_IHD")
 
 sumset <- create_summaryset(data, build = "GRCh37")
 getMetadata(sumset)

 ```
 Then we perform the liftover. We can first download the chain file followed by the `liftover()`. In this case we need to give the genome assemblies to "lift from" and to "lift to" : 

```{r}
sumset_lift <- download_chainfile(from = "GRCh37", to = "GRCh38") %>%
                liftover(sumset, chainfile = ., to = "GRCh38")
```
Or simply provide directly to `liftover()` the genome assembly to "lift to". In this case, the function will check the `build` to "lift from" in the metadata of `sumset` and download the chain file.


```{r}
sumset1_lift <- liftover(sumset, to = "GRCh38")
```

