---
title: "Creating a DataSet"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Creating a DataSet}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```
The **gwasglue2** `DataSet-class()` object contains two or more harmonised `SummarySet` objects, each one representing a GWAS study. The `DataSet` can be created from a list of `SummarySet` objects.


<br>
<br>


## Setup
In addition to **gwasglue2**, we will load the following package:

- [**ieugwasr**](https://mrcieu.github.io/ieugwasr/) - R package for accessing IEU GWAS database

```{r setup}
library(gwasglue2)
library(ieugwasr)
```


<br>
<br>

## Function
* `create_dataset()`: creates a `DataSet` object 


<br>
<br>

## Example

We will use `SummarySet` examples from the the [Input section](input.html). 

```{r}
summarysets <- list(sumset_opengwas, sumset_vcf)
```
and then use the **gwasglue2** constructor `create_dataset()` to build the `DataSet` object. By default, the constructor will harmonise the `DataSet`. More details on harmonisation can be found in the [Processing section](processing.html) or in the [Strategy](Strategy.html) page.

```{r}
dataset <-  create_dataset(summarysets)
dataset
```
If we call `dataset`, we can see summary of the `DataSet` object. For example, it shows that it contains two `SummarySet` objects, one for each study. 



