---
title: "How to create a SummarySet and a DataSet"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to create a SummarySet and a DataSet}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
   eval = FALSE
)
```

```{r setup}
library(gwasglue2)
devtools::load_all("../") # this was added just for development
```

In this tutorial, we will appproach the several ways of creating the two types of gwasglue2 objects: the `SummarySet` and the `DataSet`. Check the [Strategy](articles/Strategy.html) page for more details on these objects.


# SummarySet
 To create a SummarySet object we first need GWAS summary data, like the ones that can be obtained from IEU OpenGWAS through the `ieugwasr` package.

 ```{r}
 data1 <- ieugwasr::associations(variants = "5:74132993-75132993", id = "ukb-d-I9_IHD")
 ```
 
 Then we use the [`create_summaryset`](reference/create_summaryset.html) to create the [`SummarySet`](reference/SummarySet-class.html).

```{r}
 sumset1 <- create_summaryset(data1)
 getMetadata(sumset1)
```

In the code above, we did not supply any metadata information. The function will fill the metadata using any information avaiolable in the GWAS summary data (in `data1`).
It is possible to add metadata using the `metadata`  argument.

To create a `metadata` object, we use the [`create_metadata`](reference/create_metadata.html) function. We can give metadata information manually using one or more argument of the function. For example: 
 
```{r}
 meta1 <- create_metadata(sample_size = 361194)
 print(meta1)
```

It is also possible for the user to add extra metadata information.

```{r}
 meta1 <- create_metadata(sample_size = 361194, access = date())
 print(meta1)
```


Or to provide a dataframe with metadata information, using the `metadata` argument.

```{r}
 meta1 <- create_metadata(metadata = ieugwasr::gwasinfo("ukb-d-I9_IHD"), access = date())
 meta1  
```

Thus, we can now build a `SummarySet` object with metadata information provided. Note that `data1` is a tibble and by default `create_summaryset` reads tibbles. 

```{r}
 sumset1 <- create_summaryset(data1, metadata=meta1)
 getMetadata(sumset1)
```

Let's try with another example, using another IEU id.

First the GWAS summary data,

```{r}
 data2 <- ieugwasr::associations(variants = "5:74132993-75132993", id = "finn-b-I9_CHD")
``` 

and then the metadata

```{r}
 meta2 <- create_metadata(metadata = ieugwasr::gwasinfo("finn-b-I9_CHD"))
 meta2  
```
You will notice that there is no `sample_size` information in the metadata. This absence could cause gwasglue2 further down in the analyses. We could the  `sample_size` parameter in `create_metadata` to add the argument or add it later. We are going to add this information after creating the `SummarySet`.


```{r}
sumset2 <- create_summaryset(data2, metadata=meta2)
```

The warning tell us that there is a problem with the GWAS summary data. Specifically, there are variants with the same chromosomal position and alleles that have different betas, p-values and allele frequencies. If we change the quality control parameter to `qc = TRUE`, we will let gwasglue2 to deal with this inconsistencies in the dataset.

```{r}
sumset2 <- create_summaryset(data2, metadata=meta2, qc = TRUE)
```

Now we will deal with the lack of sample size information in the metadata. If this information was present in `data2`, the `create_summaryset` function would fill this information automatically. But this is not the case, and thus we are going to use the [`addToMetadata`](reference/addTometadata.html) method and the `ncontrol` information already present in the metadata.


```{r}
getMetadata(sumset2)
sumset2 = addToMetadata(sumset2, sample_size=getMetadata(sumset2)$ncontrol)
getMetadata(sumset2)
```

# DataSet
A [`DataSet`](reference/DataSet-class.html) is a list of two or more harmonised `SummarySets`.

First we need to create a list with all `SummarySets` we want to analyse,

```{r}
summarysets <- list(sumset1, sumset2)
```
and then use the [`create_dataset`](reference/create_dataset.html) function to build the `DataSet` object.

```{r}
dataset <-  create_dataset(summarysets, harmonise = TRUE, tolerance = 0.08, action = 1)
```



