---
title: "YAML tutorials"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{YAML tutorials}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(gwasglue2)
```

Here we demonstarte how to use the constructores to create `DataSets` for different types of analyses.




# MR analyses
 We will use the YAML constructor `job_mr()` to create a yaml object. This object will be used as input to the `gwasglue()` function to create a `DataSet`. The `job_mr()` constructor takes the following arguments:

```{r}

job_mr(exposure =("ieugwasr:ieu-a-2"),outcome = c("ieugwasr:ieu-a-7"), tophits=TRUE)

```


# A simple colocalisation analysis
The following example demonstrates how to to create  the `DataSet` for a colocalisation analysis using the `job_coloc()` constructor. 


```{r}
bed_ref <- "data/ld/EUR"

yaml <- job_coloc(traits = c("opengwas:ieu-a-2", "opengwas:ieu-a-7"), region=c("1:400000-4500000"), bfile= bed_ref, plink_bin="plink")
cat(yaml)
dataset <- gwasglue(yaml)
```




# LD score regression